---
title: "TURF Data Validation"
format:
  revealjs:
    transition: none
    width: 1050
    height: 700
    margin: 0.1
    slide-number: true
execute:
  echo: false
  warning: false
editor: source
---

```{r}
#| include: false

setwd("../..")

validation <- readRDS(
  "data/processed/turf_metrics/data_validation_objects.rds"
)

list2env(validation, envir = .GlobalEnv)
```

## how are the TURFs structured?

:::::: structure-slide
The dataset contains **169 TURF IDs** but **212 spatial sub-IDs**

::::: structure-columns
::: structure-column
**Most are simple**

165 TURF IDs → **1 sub-ID**
:::

::: structure-column
**A few contain multiple spatial areas**

MX_C_012 → **2 sub-IDs**\
MX_C_002 → **3 sub-IDs**\
MX_C_052 → **5 sub-IDs**\
MX_C_007 → **20 sub-IDs**
:::
:::::

<br>

[**→ I use sub-IDs because each sub-ID represents a separate spatial area.**]{.blue}
::::::

## which TURFs are included?

:::::::: included-slide

::: flow-row
[212]{.flow-number} <strong>TURF sub-IDs</strong>
:::

::: down-arrow
↓
:::

::: flow-row
[198]{.flow-number} <strong>have identified species</strong>
:::

::: flow-note
14 have no species information
:::

::: down-arrow
↓
:::

::: estuary-row
::: flow-row
[54]{.flow-number} <strong>have marine (non-estuarine) species</strong>
:::

::: flow-note
<em>Penaeus</em> + <em>Callinectes</em>
:::
:::

::: down-arrow
↓
:::

::: flow-row
[36]{.flow-number} <strong>have species with SDMs</strong>
:::

::::::::

## what is missing from AquaX?

::::::: filter-slide

Of the **198 sub-IDs with identified species**, only **36** contain a species with one of our processed SDMs.

:::::: species-columns

::: species-column
**Not in AquaX database**

*Gelidium*\
*Gelidium robustum*\
<span style="color:#d98989;"><em>Penaeus aztecus</em></span>\
<span style="color:#d98989;"><em>Penaeus vannamei</em></span>\
<span style="color:#d98989;"><em>Penaeus</em> spp.</span>\
<span style="color:#d98989;"><em>Callinectes</em> spp.</span>
:::

::: species-column
**In AquaX, but no SDM available**

*Anadara tuberculosa*\
<span style="color:#d98989;"><em>Callinectes rathbunae</em></span>\
<span style="color:#d98989;"><em>Penaeus stylirostris</em></span>\
*Tivela stultorum*
:::

::: species-column
**SDM potentially available**

*Octopus vulgaris*
:::

::::::

:::::::

## what do the 36 TURFs include?

::::::: modeled-species-slide
**36 TURF sub-IDs** cover [**20 species with SDMs**]{.blue}

:::::: species-columns
::: species-column
**Lobsters**

*Panulirus argus*\
*P. gracilis*\
*P. inflatus*\
*P. interruptus*\
*P. penicillatus*

**Sea urchins**

*Mesocentrotus franciscanus*\
*Strongylocentrotus purpuratus*
:::

::: species-column
**Abalone**

*Haliotis corrugata*\
*H. cracherodii*\
*H. fulgens*\
*H. rufescens*\
*H. sorenseni*

**Oysters**

*Crassostrea virginica*\
*Striostrea prismatica*
:::

::: species-column
**Sea cucumbers**

*Apostichopus parvimensis*\
*Holothuria inhabilis*\
*H. impatiens*

**Other**

*Megastraea undosa*\
*Melongena patula*\
*Cardisoma guanhumi*
:::
::::::
:::::::

## where are the 36 modeled sub-IDs?

:::::: map-slide

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 5.4
#| out-width: "100%"
#| fig-align: center

modeled_turfs_plot
```

::::::


## how many AquaX cells support each observation?

:::::: cell-support-slide

<div class="cell-support-intro">

Each observation = **1 sub-ID × 1 species × 1 scenario**  
**749 observations total**

</div>

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9
#| fig-height: 3.7
#| out-width: "72%"
#| fig-align: center

cell_support_plot
```

<div class="cell-support-takeaway">

**222 of 749 observations (29.6%)** have ≤3 HSI-valued cells.  
Of these, **52 observations (6.9%)** have **only NA-valued cells**.

</div>

::::::
## what did we find about AquaX NAs?

:::::: na-slide

### 1. some sub-IDs have no AquaX data

**5 sub-IDs** are completely NA for ≥1 species  
*Crassostrea virginica · Strongylocentrotus purpuratus · Mesocentrotus franciscanus*

*→ all NA from current through 2100*

### 2. two cases actually lose their data

**2 sub-ID × species combinations** become NA in future scenarios  
*S. purpuratus*

→ all NA after RCP2.6 2050

### 3. partial NAs are widespread, but stable

**45 sub-ID × species combinations**

Same number of NA cells across **all scenarios and time periods**

`→ If NA = no prediction: partial NAs are less concerning because they are stable and excluded from calculations; all-NA cases cannot be used to calculate metrics or interpreted as habitat absence.`

::::::

## MEAN HSI · DATA CHECK 

<div class="mean-hsi-na-slide">

### **107 observations currently have mean HSI = NA**

<br>

**55 observations** have HSI-valued cells, but **no cells above the suitability cutoff**

`→ No suitable habitat → mean HSI = 0`

<br>

**52 observations** have **only NA-valued cells**

`→ Mean HSI cannot be calculated → remains NA`

<br>

<div class="metric-check-takeaway">

**→ We need to distinguish absence of suitable habitat (0) from absence of an AquaX prediction (NA).**

</div>

</div>

## MEAN HSI · WHERE ARE THE 55 ZERO-SUITABILITY CASES?

<div class="mean-hsi-attention-slide">

```{r}
#| echo: false
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "92%"
#| fig-align: center

mean_hsi_zero_map
```

<div class="attention-takeaway">

**→ These represent no suitable habitat and should have mean HSI = 0, rather than NA.**

</div>

</div>

## MEAN HSI · WHERE ARE THE 52 NA CASES

<div class="mean-hsi-attention-slide">

<div class="attention-map-columns">

<div class="attention-map">

**Baja California**

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4.5
#| out-width: "100%"

mean_hsi_attention_bc_plot
```

</div>

<div class="attention-map">

**Veracruz**

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 5
#| fig-height: 4.5
#| out-width: "100%"

mean_hsi_attention_ver_plot
```

</div>

</div>

<div class="attention-legend">

<span class="legend-item">
<span class="legend-box permanent-na"></span>
**All NA in every scenario**
</span>

<span class="legend-item">
<span class="legend-box future-na"></span>
**Becomes all NA in some scenarios**
</span>

<span class="legend-item">
<span class="legend-box no-issue"></span>
Other modeled sub-IDs
</span>

</div>

</div>

## MEAN HSI · CLIMATE RESPONSE

:::::: mean-hsi-na-slide

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 4.5
#| out-width: "70%"
#| fig-align: center

mean_hsi_climate_plot
```

<div class="percent-bottom-text">

Average mean HSI declines from **830 under current conditions** to **753 under RCP8.5 2100**.

**RCP8.5 2100** has lowest average mean HSI..

</div>

::::::

## % SUITABLE HABITAT · DATA CHECK

:::::: percent-suitable-slide

### **697 of 749 observations have HSI-valued cells**

<br>

**All 697 independently reproduce the calculated % suitable**

`→ No calculation errors found`

<br>

**170 observations are calculated from only 1–3 HSI-valued cells**

`→ 1 cell: 0% or 100%`

`→ 2 cells: 0%, 50%, or 100%`

`→ 3 cells: 0%, 33.3%, 66.7%, or 100%`

<br>

**52 observations have no HSI-valued cells**

`→ % suitable cannot be calculated → remains NA`

<div class="metric-check-takeaway">

**→ The calculation is correct, but some % suitable estimates have very coarse spatial support.**

</div>

::::::


## % SUITABLE HABITAT · WHERE ARE THESE CASES?

<div class="mean-hsi-attention-slide percent-attention-slide">

<div class="attention-map-columns percent-map-columns">

<div class="attention-map">

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 5

percent_suitable_attention_west_plot
```

</div>

<div class="attention-map">

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 5

percent_suitable_attention_ver_plot
```

</div>

</div>

<div class="attention-legend">

<span class="legend-item">
<span class="legend-box zero-cells"></span>
**0 HSI-valued cells**
</span>

<span class="legend-item">
<span class="legend-box one-cell"></span>
**1 cell**
</span>

<span class="legend-item">
<span class="legend-box two-three-cells"></span>
**2–3 cells**
</span>

<span class="legend-item">
<span class="legend-box transition-cells"></span>
**2 cells → 0 cells**
</span>

<span class="legend-item">
<span class="legend-box no-issue"></span>
Other modeled sub-IDs
</span>

</div>

<div class="attention-takeaway">

**→ These sub-IDs need extra attention because % suitable is either unavailable or based on very few HSI-valued cells.**

</div>

</div>

## % SUITABLE HABITAT · CLIMATE RESPONSE

<div class="percent-suitable-slide" data-auto-animate="false">

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 4.5
#| out-width: "70%"
#| fig-align: center

percent_100_plot
```

<div class="percent-bottom-text">

**60 of 107 combinations (56.1%)** are 100% suitable under current conditions.

By **RCP8.5 2100**, this falls to **38 of 107 (35.5%)**.

</div>

</div>

## % SUITABLE HABITAT · REDISTRIBUTION

<div class="percent-suitable-slide">

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 4.5
#| out-width: "78%"
#| fig-align: center

percent_suitable_redistribution_plot
```

<div class="percent-bottom-text">

Under **RCP8.5 2100**, across 107 TURF × species combinations:

**11 gain** · **38 lose** · **50 show no change** · **8 cannot be compared**

Changes range from **−100 to +100 percentage points**.

*All ±100 percentage-point changes occur in sub-ID 29_1 and are based on 2 HSI-valued cells.*

</div>

</div>


## % SUITABLE HABITAT · WHERE DO RESPONSES OCCUR?

<div class="percent-suitable-slide">

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "88%"
#| fig-align: center

percent_suitable_direction_map
```

<div class="metric-check-takeaway">

**→ Some sub-IDs have species that gain suitable habitat and species that lose suitable habitat.**

</div>

</div>

## SPECIES RICHNESS · DATA CHECK

<div class="percent-suitable-slide">

### **Tau (threshold) changed from >10% to >0% suitable habitat**

<br>

Species is now counted whenever any suitable habitat is present.

<br>

**239 of 252 observations (94.8%)** are unchanged.

**13 of 252 observations (5.2%)** increase by exactly **1 species**.

**12 of those 13 differences involve *Panulirus argus*.**

<div class="metric-check-takeaway">

**→ Removing the 10% threshold has little effect on species richness.**

</div>

</div>

## SPECIES RICHNESS · CLIMATE RESPONSE

<div class="mean-hsi-attention-slide">

```{r}
#| echo: false
#| fig-width: 11
#| fig-height: 5.0
#| out-width: "92%"
#| fig-align: center

richness_response_map
```

<div class="attention-takeaway">

Under **RCP8.5 in 2100**, **4 of 36 sub-IDs show a modeled loss in target-species richness**.

**2 additional apparent losses** occur because the future AquaX prediction becomes NA, while **2 apparent gains** involve species already targeted by the TURF that have 0% suitable habitat currently.

</div>

</div>

## POTENTIAL REDISTRIBUTION · GAINS & LOSSES

<div class="mean-hsi-na-slide">

**Metric 4 tests all 20 modeled species across all 36 modeled sub-IDs, to identify potential gains from species that a TURF does not currently target.**

<br>

| Change under RCP8.5 2100 | Number | What it means |
|---|---:|---|
| **Target species lost** | **6** | Current target goes from >0% → 0% suitable |
| **Existing target becomes suitable** | **3** | Current target goes from 0% → >0% suitable |
| **New non-target species becomes suitable** | **0** | Species not currently targeted goes from 0% → >0% suitable |
| **Target species becomes NA** | **2** | Future AquaX prediction becomes unavailable |

<br>

<div class="metric-check-takeaway">

**→ Expanding the analysis to all modeled species identified no potential gains from species outside a TURF's current target list.**

</div>

</div>


<style>

/* -------------------------------------------------------


   GENERAL
------------------------------------------------------- */

.reveal h2 {
  color: #1769aa;
  font-size: 1.25em;
  font-weight: 700;
  text-transform: none;
  margin-bottom: 0.7em;
}

.blue {
  color: #1769aa;
}


/* -------------------------------------------------------
   SLIDE 1: TURF STRUCTURE
------------------------------------------------------- */

.structure-slide {
  font-size: 0.90em;
}

.structure-columns {
  margin-top: 0.7em;
  display: grid;
  grid-template-columns: 1fr 1fr;
  column-gap: 3em;
  align-items: start;
}

.structure-column {
  width: 100%;
}


/* -------------------------------------------------------
   SLIDE 2: WHICH TURFS ARE INCLUDED?
------------------------------------------------------- */

.included-slide {
  font-size: 0.78em;
  line-height: 1.15;
}

.flow-row {
  font-size: 1em;
  margin: 0;
}

.flow-number {
  color: #1769aa;
  font-size: 1.35em;
  font-weight: 700;
}

.down-arrow {
  color: #1769aa;
  font-size: 1.35em;
  line-height: 1;
  margin: 0.25em 0;
}

.flow-note {
  color: #777;
  font-size: 0.72em;
  font-style: italic;
  margin-top: 0.05em;
}


/* -------------------------------------------------------
   SLIDE 3: WHY ARE SUB-IDs NOT INCLUDED?
------------------------------------------------------- */

.filter-slide {
  font-size: 0.82em;
}

.filter-takeaway {
  color: #1769aa;
  font-size: 0.70em;
  line-height: 1.2;
  margin-top: 1em;
}


/* -------------------------------------------------------
   SLIDE 4: MODELED SPECIES
------------------------------------------------------- */

.modeled-species-slide {
  font-size: 0.88em;
}


/* -------------------------------------------------------
   THREE-COLUMN SPECIES LAYOUT
------------------------------------------------------- */

.species-columns {
  margin-top: 0.7em;
  display: grid;
  grid-template-columns: repeat(3, 1fr);
  column-gap: 2em;
  align-items: start;
}

.species-column {
  width: 100%;
  padding: 0;
  margin: 0;
}

.species-column p {
  font-size: 0.70em;
  line-height: 1.3;
  margin-top: 0.15em;
  margin-bottom: 0.15em;
  white-space: nowrap;
}

.species-column p:has(strong) {
  font-size: 0.88em;
  margin-top: 0.65em;
  margin-bottom: 0.35em;
  line-height: 1.1;
}

.species-column p:has(strong) strong {
  color: #1769aa;
  font-weight: 700;
}


/* -------------------------------------------------------
   SLIDE 5: MODELED TURF MAP
------------------------------------------------------- */

.map-slide {
  font-size: 0.88em;
}


/* -------------------------------------------------------
   SLIDE 6: AQUAX CELL SUPPORT
------------------------------------------------------- */

.cell-support-slide {
  font-size: 0.56em;
}

.cell-support-intro {
  font-size: 0.9em;
  line-height: 1.15;
  margin-top: 0.1em;
  margin-bottom: 0.1em;
}

.cell-support-takeaway {
  color: #1769aa;
  font-size: 0.9em;
  line-height: 1.15;
  text-align: center;
  margin-top: 0;
}


/* -------------------------------------------------------
   SLIDE 7: AQUAX NA PATTERNS
------------------------------------------------------- */

.na-slide {
  font-size: 0.60em;
  line-height: 1.18;
}

.na-slide table {
  font-size: 0.95em;
  margin-top: 0.35em;
  margin-bottom: 0.65em;
}

.na-slide p {
  margin-top: 0.4em;
  margin-bottom: 0.4em;
}

.na-takeaway {
  color: #1769aa;
  font-size: 0.95em;
  line-height: 1.18;
  margin-top: 0.6em;
}


/* -------------------------------------------------------
   METRIC: MEAN HSI
------------------------------------------------------- */

.mean-hsi-na-slide {
  font-size: 0.60em;
  line-height: 1.15;
}

.mean-hsi-na-slide h3 {
  font-size: 1.15em;
  margin-top: 0.7em;
  margin-bottom: 0.4em;
}

/* -------------------------------------------------------
   MEAN HSI: CASES NEEDING EXTRA ATTENTION
------------------------------------------------------- */

.mean-hsi-attention-slide {
  font-size: 0.60em;
  line-height: 1.15;
}

.attention-map-columns {
  display: grid;
  grid-template-columns: 1fr 1fr;
  column-gap: 1.5em;
  align-items: start;
}

.attention-map {
  text-align: center;
}

.attention-map > p:first-child {
  color: #1769aa;
  font-size: 1.05em;
  margin-top: 0;
  margin-bottom: 0.15em;
}

.attention-legend {
  display: flex;
  justify-content: center;
  align-items: center;
  gap: 2em;
  margin-top: 0.25em;
  font-size: 0.82em;
}

.legend-item {
  display: inline-flex;
  align-items: center;
  gap: 0.4em;
}

.legend-box {
  display: inline-block;
  width: 1.1em;
  height: 1.1em;
  border: 1px solid black;
}

.permanent-na {
  background-color: #d95f5f;
}

.future-na {
  background-color: #f0ad4e;
}

.no-issue {
  background-color: #cccccc;
}

.attention-takeaway {
  color: #1769aa;
  text-align: center;
  font-size: 0.90em;
  margin-top: 0.65em;
}

/* -------------------------------------------------------
   METRIC: % SUITABLE HABITAT
------------------------------------------------------- */

.percent-suitable-slide {
  font-size: 0.60em;
  line-height: 1.15;
}

.metric-label {
  color: #1769aa;
  font-size: 0.70em;
  font-weight: 700;
  letter-spacing: 0.05em;
  margin-top: 0;
  margin-bottom: 0.15em;
}

.percent-suitable-takeaway {
  line-height: 1.15;
  margin-top: 0.25em;
}

.metric-check-takeaway {
  color: #1769aa;
  font-size: 0.95em;
  line-height: 1.18;
  margin-top: 0.5em;
}

.percent-bottom-text {
  font-size: 0.75em;
  line-height: 1.15;
  margin-top: 0.2em;
}

/* % suitable diagnostic-map legend */

.zero-cells {
  background-color: #d95f5f;
}

.one-cell {
  background-color: #f0ad4e;
}

.two-three-cells {
  background-color: #f6d55c;
}

.transition-cells {
  background-color: #9b59b6;
}

</style>